---
title: "1_caro_spect_data_loadings"
output: html_document
date: "2024-01-16"
---

This data is obtained from the eccentric and concentric exercise trained rats to explore the *training-induced adaptations*. Using LCM, we specifically collected **MTJ, only tendon and only muscle**. 
Database search is done by spectronaut 5/Jan/2024. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T)

source(here::here("R/library.R"))
rm(list = ls())
```

# load metadata
<!-- make sure metadata has the same sample IDs and the same sample order -->
```{r}
metadata <- read.csv2("data-raw/MetaData_10Jan.csv", header = T, sep = ';') %>% 
    dplyr::select(1:15) %>%
    mutate_at(1:15, as.factor) %>% 
    arrange(tissue)

usethis::use_data(metadata, overwrite = T)
```


# load data
gene names are used as rownames. if not present, they were replaced by protein names. 
log2 transformation

Mistake with sample ID: 94 is written twice in the samples_export file provided to PRI. One of them is 94 and the other one is 84, but we cannot find out which is which now. Since 84 belongs to SED and 94 belongs to ECC groups, we might need to have an idea based on protein expression.
```{r}
df <- read.delim("data-raw/20240109_093728_20231215_ProteinQuantReport.tsv", header = T, check.names = T, sep = "\t") %>%
        dplyr::select("PG.ProteinGroups", "PG.Genes", 6:113)

new_colnames <- sapply(strsplit(colnames(df), "_"), '[',10)
colnames(df) <- new_colnames

colnames(df)[1] <- "protein_names"
colnames(df)[2] <- "genes"
colnames(df)[96] <- "S109" #mistake in the samples_export list S94 is written twice for plate position H2 and H10, so S84 is missing in the list. Either of these S94 is: S94 ECC 5d muscle or S84 sed 5b muscle! Keep them for now. 

counts <- table(df$genes)

df <- df %>%
    mutate(genes = if_else(genes == "", protein_names, genes)) %>%
    mutate_at(3:110, as.numeric) %>%
    mutate_all(~ifelse(is.nan(.), NA, .)) %>%
    mutate(genes = make.names(genes, unique = T), genes) %>% 
    column_to_rownames(var = "genes") %>% 
    dplyr::select(-c("protein_names")) %>%
    log2()

names(df) <- sub("^S", "", names(df))

#order df by tissue to make sure sample order in the main data matrix follows the metadata!!
sample_order <- metadata$sample_id[order(metadata$tissue)]
sample_order <- trimws(sample_order) 
df <- df[ ,sample_order]

usethis::use_data(df, overwrite = T)
write.csv(df, file = "data/caro_ex_lcm_spect_allsamples_sorted.csv", row.names = T, quote = F)
```

# muscle only 
1798 proteins 
```{r}
df_muscle <- read.delim("data-raw/20240225_194041_Caro-LCM-onlymuscle_proteinquant_Report.tsv", header = T, check.names = T, sep = "\t") %>%
        dplyr::select("PG.ProteinGroups", "PG.Genes", 5:40) 

new_colnames <- sapply(strsplit(colnames(df_muscle), "_"), '[',10)
colnames(df_muscle) <- new_colnames

colnames(df_muscle)[1] <- "protein_groups"
colnames(df_muscle)[2] <- "genes"
colnames(df_muscle)[33] <- "S109"

df_muscle <- df_muscle %>%
    mutate(genes = if_else(genes == "", protein_groups, genes)) %>%
    mutate_at(3:38, as.numeric) %>%
    mutate_all(~ifelse(is.nan(.), NA, .)) %>%
    mutate(genes = make.names(genes, unique = T), genes) %>% 
    column_to_rownames(var = "genes") %>% 
    dplyr::select(-c("protein_groups")) %>%
    log2()

names(df_muscle) <- sub("^S", "", names(df_muscle))

#############
# metadata, sample order 
metadata <- read.csv2("data-raw/MetaData_10Jan.csv", header = T, sep = ';') %>% 
    dplyr::select(1:15) %>%
    mutate_at(1:15, as.factor) %>% 
    arrange(intervention)

metadata_muscle <- metadata[metadata$tissue == "Muscle", ]

sample_order <- metadata_muscle$sample_id[order(metadata_muscle$intervention)]
sample_order <- trimws(sample_order) 
df_muscle <- df_muscle[ ,sample_order]

###############
# data vis
vis_miss(df_muscle, sort_miss = T)
plotDensities(df_muscle, main = "caro_ex_lcm_onlymuscle_log2")
boxplot(df_muscle, main = "caro_ex_lcm_onlymuscle_log2")

df_muscle_cleaned <- df_muscle %>% dplyr::select(-"19")
plotDensities(df_muscle_cleaned, main = "caro_ex_lcm_onlymuscle_log2_cleaned")
boxplot(df_muscle_cleaned, main = "caro_ex_lcm_onlymuscle_log2_cleaned")

metadata_muscle <- metadata_muscle %>% filter(sample_id != "19" )

################
# 50% filter > most abundant proteins 

df_muscle_cleaned50 <- selectOverallPercent(df_muscle_cleaned, 0.5)
df_muscle_cleaned50$median <- rowMedians(as.matrix(df_muscle_cleaned50), na.rm = T)
df_muscle_cleaned50 <- df_muscle_cleaned50 %>% arrange(desc(median))
df_muscle_cleaned50$rank = (1:1667) 
df_muscle_cleaned50$protein <- row.names(df_muscle_cleaned50)


################
# 50% filter > proteins from the literature 
known_mtj_prot <- c("Xirp2", "Thbs4", "Ncam1", "Col22a1", "Fermt2", "Cd151", "Cilp", "Itga10", "Ogn", "Fmod", "Bgn", "Col1a2", "Col6a3", "Col6a2", "Col1a1", "Col6a1", "Col12a1", "Prelp", "Lam2", "Vdaxc1", "Vdac2", "Vdac3", "Mpz", "Vcl", "Plec", "Des", "Dsp", "Actn" , "Capz", "Flnc", "Fhl1", "Ttn", "Pdlim7", "Myot")

poi_indata <- df_muscle_cleaned[row.names(df_muscle_cleaned) %in% known_mtj_prot, ]

poi_indata_long <- poi_indata %>% 
    tibble::rownames_to_column("genes") %>% 
    tidyr::pivot_longer(
        cols = !genes,
        values_to = "intensities",
        names_to = "sample_id"
    ) %>% 
    dplyr::inner_join(metadata_muscle) %>% 
    dplyr::select(genes, sample_id, intensities, tissue, intervention)

head(poi_indata_long)

ggplot(poi_indata_long, aes(x = tissue, y = intensities)) + 
  geom_violin(trim = FALSE) +  # Set trim to FALSE to display the entire violin plot
  geom_point(position = position_jitter(width = 0.2), size = 1.5, alpha = 0.5) +
  stat_summary(fun = median, geom = "crossbar", width = 0.5, color = "black") +
  labs(x = "Tissue", y = "log2 protein intensities", title = "known_mtj_proteins_onlymuscledata") +
  theme_minimal() + 
  facet_wrap(~ genes, scales = "free")


################
known_tendon_prot <- c("Serpina1", "Dcn", "Fbn1", 
"Fn1", "Col1a1", "Col1a2", "Col3a1", "Col4a1", "Col5a1", "Col5a2", 
"Col6a1", "Col6a2", "Col6a3", "Col11a1", "Col11a2", "Col2a1", "Col12a1", "Comp")

poi_indata <- df_muscle_cleaned[row.names(df_muscle_cleaned) %in% known_tendon_prot, ]

poi_indata_long <- poi_indata %>% 
    tibble::rownames_to_column("genes") %>% 
    tidyr::pivot_longer(
        cols = !genes,
        values_to = "intensities",
        names_to = "sample_id"
    ) %>% 
    dplyr::inner_join(metadata_muscle) %>% 
    dplyr::select(genes, sample_id, intensities, tissue, intervention)

head(poi_indata_long)

ggplot(poi_indata_long, aes(x = tissue, y = intensities)) + 
  geom_violin(trim = FALSE) +  # Set trim to FALSE to display the entire violin plot
  geom_point(position = position_jitter(width = 0.2), size = 2, alpha = 0.5) +
  stat_summary(fun = median, geom = "crossbar", width = 0.5, color = "black") +
  labs(x = "Tissue", y = "log2 protein intensities", title = "known_tendon_proteins_onlymuscledata") +
  theme_minimal() + 
  facet_wrap(~ genes, scales = "free")


```

# summary reports
```{r}
summary(df)
#str(df) #all numeric
```


# vis number of NA
done by vis_dat and vis_miss functions.
```{r}
vis_dat(df)
vis_miss(df, sort_miss = T)
```

# vis number of NA vs number of valid values per sample
```{r}
counts_na <- sapply(df, function (x) sum(is.na(x)))
counts_valid <- sapply(df, function (x) sum(!is.na(x)))

df_na <- tibble(counts_na, counts_valid)
df_na$tissue <- metadata$tissue

df_na <- df_na %>% 
    arrange(tissue) #first, need to sort the sample order based on tissue, and then add sample_id info

df_na$sampleid <- metadata$sample_id

df_na <- df_na %>% mutate(sampleid = factor(sampleid, levels = sampleid))

ggplot(df_na, aes(x = sampleid)) + 
    geom_bar(aes(y = counts_valid, fill = tissue), stat = "identity", position = "dodge", alpha = 0.5) +
  geom_bar(aes(y = counts_na, fill = tissue), stat = "identity", position = "dodge", alpha = 0.9) +
  labs(title = "counts_valid vs counts_na per sample spect", y = "counts") + 
  theme_minimal()

#ggsave(file = "doc/caro_ex_lcm_spect_allsamples_raw_counts_valid_vs_counts_na.jpeg", width=10, height = 8)

```

# any samples to remove
sample_id 65 has almost 50% NA - is removed from the data.
```{r}
plotDensities(df, main = "caro_ex_lcm_allsamples_log2")

df_cleaned <- df %>% dplyr::select(-"65")

plotDensities(df_cleaned, main = "caro_ex_lcm_1sampleremoved_log2")

usethis::use_data(df_cleaned, overwrite = T)
write.csv(df_cleaned, file = "data/caro_ex_lcm_spect_cleaned.csv", row.names = T, quote = F)

metadata_cleaned <- metadata %>% filter(sample_id != "65")
usethis::use_data(metadata_cleaned, overwrite = T)
```

