---
title: "1_caro_spect_data_loadings"
output: html_document
date: "2024-01-16"
---

This data is obtained from the eccentric and concentric exercise trained rats to explore the *training-induced adaptations*. Using LCM, we specifically collected **MTJ, only tendon and only muscle**. 
Database search is done by spectronaut 5/Jan/2024. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T)

source(here::here("R/library.R"))
rm(list = ls())
```

#load metadata
<!-- make sure metadata has the same sample IDs and the same sample order -->
```{r}
metadata <- read.csv2("data-raw/MetaData_10Jan.csv", header = T, sep = ';') %>% 
    dplyr::select(1:15) %>%
    mutate_at(1:15, as.factor) %>% 
    arrange(tissue)

usethis::use_data(metadata, overwrite = T)
```


#load data
gene names are used as rownames. if not present, they were replaced by protein names. 
log2 transformation

Mistake with sample ID: 94 is written twice in the samples_export file provided to PRI. One of them is 94 and the other one is 84, but we cannot find out which is which now. Since 84 belongs to SED and 94 belongs to ECC groups, we might need to have an idea based on protein expression.
```{r}
df <- read.delim("data-raw/20240109_093728_20231215_ProteinQuantReport.tsv", header = T, check.names = T, sep = "\t") %>%
        dplyr::select("PG.ProteinGroups", "PG.Genes", 6:113)

new_colnames <- sapply(strsplit(colnames(df), "_"), '[',10)
colnames(df) <- new_colnames

colnames(df)[1] <- "protein_names"
colnames(df)[2] <- "genes"
colnames(df)[96] <- "S109" #mistake in the samples_export list S94 is written twice for plate position H2 and H10, so S84 is missing in the list. Either of these S94 is: S94 ECC 5d muscle or S84 sed 5b muscle! Keep them for now. 

counts <- table(df$genes)

df <- df %>%
    mutate(genes = if_else(genes == "", protein_names, genes)) %>%
    mutate_at(3:110, as.numeric) %>%
    mutate_all(~ifelse(is.nan(.), NA, .)) %>%
    mutate(genes = make.names(genes, unique = T), genes) %>% 
    column_to_rownames(var = "genes") %>% 
    dplyr::select(-c("protein_names")) %>%
    log2()

names(df) <- sub("^S", "", names(df))

#order df by tissue to make sure sample order in the main data matrix follows the metadata!!
sample_order <- metadata$sample_id[order(metadata$tissue)]
sample_order <- trimws(sample_order) 
df <- df[ ,sample_order]

usethis::use_data(df, overwrite = T)
write.csv(df, file = "data/caro_ex_lcm_spect_allsamples_sorted.csv", row.names = T, quote = F)
```

#summary reports
```{r}
summary(df)
#str(df) #all numeric
```


#visualization of number of NA
done by vis_dat and vis_miss functions.
```{r}
vis_dat(df)
vis_miss(df, sort_miss = T)
```

#visualization of number of NA vs number of valid values per sample
```{r}
counts_na <- sapply(df, function (x) sum(is.na(x)))
counts_valid <- sapply(df, function (x) sum(!is.na(x)))

df_na <- tibble(counts_na, counts_valid)
df_na$tissue <- metadata$tissue

df_na <- df_na %>% 
    arrange(tissue) #first, need to sort the sample order based on tissue, and then add sample_id info

df_na$sampleid <- metadata$sample_id

df_na <- df_na %>% mutate(sampleid = factor(sampleid, levels = sampleid))

ggplot(df_na, aes(x = sampleid)) + 
    geom_bar(aes(y = counts_valid, fill = tissue), stat = "identity", position = "dodge", alpha = 0.5) +
  geom_bar(aes(y = counts_na, fill = tissue), stat = "identity", position = "dodge", alpha = 0.9) +
  labs(title = "counts_valid vs counts_na per sample spect", y = "counts") + 
  theme_minimal()

#ggsave(file = "doc/caro_ex_lcm_spect_allsamples_raw_counts_valid_vs_counts_na.jpeg", width=10, height = 8)

```


#any samples to remove
sample_id 65 has almost 50% NA - is removed from the data.
```{r}
plotDensities(df, main = "caro_ex_lcm_allsamples_log2")

df_cleaned <- df %>% dplyr::select(-"65")

plotDensities(df_cleaned, main = "caro_ex_lcm_1sampleremoved_log2")

usethis::use_data(df_cleaned, overwrite = T)
write.csv(df_cleaned, file = "data/caro_ex_lcm_spect_cleaned.csv", row.names = T, quote = F)

metadata_cleaned <- metadata %>% filter(sample_id != "65")
usethis::use_data(metadata_cleaned, overwrite = T)
```



#NA filter whole matrix and normalize
raw data matrix is saved as Rda file. Similarly, data is saved separately after NA filtering.
```{r}
df_70f <- selectOverallPercent(df_cleaned, 0.7) #1784 proteins << original df was 2318

df_70f_qn <- normalizeQuantiles(df_70f)

usethis::use_data(df_70f_qn, overwrite = T)
write.csv(df_70f_qn, file = "data/caro_ex_lcm_spect_1sampleremoved_sorted_70f_qn.csv", row.names = T, quote = F)
```


#data visualization
```{r}
metadata.grp.col <- c("#F8766D", "#00BA38", "#619CFF")
grp.col <- metadata.grp.col[match(metadata_cleaned$tissue, unique(metadata_cleaned$tissue))]

boxplot(df_70f_qn, main = "caro_spect_1sampleremoved_NA70f_qn", ylab = "log2 intensities", col = grp.col)
legend("topleft", inset = .02,
       c("MTJ","Muscle","Tendon"), 
       fill = metadata.grp.col, 
       horiz = TRUE, cex = 0.8)
```

# MDS plot 
***MDS shows a region-specific clustering*** There is no clustering based on intervention, duration of intervention, MS sampleprep date, cryostat date, lcm date or duration of freezing. 
```{r}
mds_70f_qn <- plotMDS(df_70f_qn)

toplot = data.frame(Dim1 = mds_70f_qn$x, Dim2 = mds_70f_qn$y, 
                    samples = metadata_cleaned$sample_id, 
                    area = metadata_cleaned$tissue, 
                    intervention = metadata_cleaned$intervention, 
                    cryo_date = metadata_cleaned$cryostat, 
                    lcm_date = metadata_cleaned$lcm, 
                    sampleprep_date = metadata_cleaned$sampleprep_date,
                    duration_frozen_slide = metadata_cleaned$duration_frozen_slide, 
                    duration_frozen_lcmtissue = metadata_cleaned$duration_frozen_microdis)

#####by area 
ggplot(toplot, aes(x = Dim1 , y = Dim2, colour = area, label = samples)) +
    geom_point(size = 3) +
    xlab(paste("Leading logFC dim1 - 63%")) + 
    ylab(paste("Leading logFC dim2 - 7%")) + 
    ggtitle("df_1sampleremoved_70fNA_matrix_qn_MDS") + 
    theme_minimal()
ggsave(file = "doc/caro_ex_lcm_spect_1sampleremoved_70f_matrix_qn_MDS_byarea.jpeg")

#####by intervention 
ggplot(toplot, aes(x = Dim1 , y = Dim2, colour = area, shape = intervention, label = samples)) +
    geom_point(size = 3) +
    scale_shape_manual(values = c("cc" = 1, "ecc" = 13, "sed" = 19)) +
    xlab(paste("Leading logFC dim1 - 63%")) + 
    ylab(paste("Leading logFC dim2 - 7%")) + 
    ggtitle("df_1sampleremoved_70fNA_matrix_qn_MDS") + 
    theme_minimal()
ggsave(file = "doc/caro_ex_lcm_spect_1sampleremoved_70f_matrix_qn_MDS_byintervention.jpeg")

#####by sampleprep_date
ggplot(toplot, aes(x = Dim1 , y = Dim2, colour = area, shape = sampleprep_date, label = samples)) +
    geom_point(size = 3) +
    scale_shape_manual(values = c("1" = 0, "2" = 15)) +
    xlab(paste("Leading logFC dim1 - 63%")) + 
    ylab(paste("Leading logFC dim2 - 7%")) + 
    ggtitle("df_1sampleremoved_70fNA_matrix_qn_MDS") + 
    theme_minimal()
ggsave(file = "doc/caro_ex_lcm_spect_1sampleremoved_70f_matrix_qn_MDS_bysampleprep.jpeg")

#####by cryo_date
ggplot(toplot, aes(x = Dim1 , y = Dim2, shape = cryo_date, colour = sampleprep_date, label = samples)) +
    geom_point(size = 3) +
    scale_shape_manual(values = c("02/Oct" = 1, "06/Oct" = 2, "11/Oct" = 3, "22/Sep" = 4, "22/Sep" = 5, "25/Oct" = 6, "25/Sep" = 7, "26/Oct" = 8, "27/Sep" = 9, "28/Oct" = 10, "28/Sep" = 11, "29/Sep" = 12)) +
    xlab(paste("Leading logFC dim1 - 63%")) + 
    ylab(paste("Leading logFC dim2 - 7%")) + 
    ggtitle("df_1sampleremoved_70fNA_matrix_qn_MDS") + 
    theme_minimal()
ggsave(file = "doc/caro_ex_lcm_spect_1sampleremoved_70f_matrix_qn_MDS_bycryo.jpeg")

#####by lcm_date
ggplot(toplot, aes(x = Dim1 , y = Dim2, shape = lcm_date, colour = sampleprep_date, label = samples)) +
    geom_point(size = 3) +
    scale_shape_manual(values = c("01/Nov" = 1, "02/Nov" = 2, "09/Oct" = 3, "10/Oct" = 4, "12/Oct" = 5, "16/Oct" = 6, "17/Oct" = 7, "18/Oct" = 8, "20/Oct" = 9, "23/Oct" = 10, "24/Oct" = 11, "26/Oct" = 12, "27/Oct" = 13, "30/Oct" = 14, "31/Oct" = 15)) +
    xlab(paste("Leading logFC dim1 - 63%")) + 
    ylab(paste("Leading logFC dim2 - 7%")) + 
    ggtitle("df_1sampleremoved_70fNA_matrix_qn_MDS") + 
    theme_minimal()
ggsave(file = "doc/caro_ex_lcm_spect_1sampleremoved_70f_matrix_qn_MDS_bylcm.jpeg")


#####by duration_frozen_days slides
ggplot(toplot, aes(x = Dim1 , y = Dim2, shape = sampleprep_date, colour = duration_frozen_slide, label = samples)) +
    geom_point(size = 3) +
    scale_shape_manual(values = c("1" = 1, "2" = 19)) +
    xlab(paste("Leading logFC dim1 - 63%")) + 
    ylab(paste("Leading logFC dim2 - 7%")) + 
    ggtitle("df_1sampleremoved_70fNA_matrix_qn_MDS") + 
    theme_minimal()
ggsave(file = "doc/caro_ex_lcm_spect_1sampleremoved_70f_matrix_qn_MDS_bydurationfrozenslide.jpeg")


#####by duration_frozen_days tissues microdissected
ggplot(toplot, aes(x = Dim1 , y = Dim2, shape = sampleprep_date, colour = duration_frozen_lcmtissue, label = samples)) +
    geom_point(size = 3) +
    scale_shape_manual(values = c("1" = 1, "2" = 19)) +
    xlab(paste("Leading logFC dim1 - 63%")) + 
    ylab(paste("Leading logFC dim2 - 7%")) + 
    ggtitle("df_1sampleremoved_70fNA_matrix_qn_MDS") + 
    theme_minimal()
ggsave(file = "doc/caro_ex_lcm_spect_1sampleremoved_70f_matrix_qn_MDS_bydurationfrozenlcmtissue.jpeg")
```

# PCA using 100%filtered data (na.omit)
Here; all NAs are omitted, resulting **810 proteins** in total. Afterwards, data is quantile normalized.

***PCA shows a region-specific clustering*** There is no clustering based on intervention, duration of intervention, MS sampleprep date, cryostat date, lcm date or duration of freezing. 
```{r}

principal_component_analysis <-
  function(.data, our_variable) {
    data_pca <- t(.data) |>
      as.matrix()
    pca_object <- stats::prcomp(data_pca, scale = T)
    data_pca <- pca_object$x
    data_pca <- as.data.frame(data_pca)
    data_pca <- data_pca |>
      dplyr::select(PC1, PC2) |>
      tibble::add_column(our_variable = our_variable)
    pca_plot <- ggplot(data_pca,
                       aes(
                         x = PC1,
                         y = PC2,
                         color = our_variable,
                         names = sample_id
                       )) +
      geom_point() +
      labs(colour = deparse(substitute(our_variable)))+
      theme_minimal() +
      theme(plot.title = element_text(size = 16, face = "bold")) +
      theme(plot.title = element_text(hjust = 0.5)) +
      theme(legend.title = element_text(size = 12, face = "bold")) +
      scale_color_viridis_d()
    return(plotly::ggplotly(pca_plot))
  }



df_50f <- selectOverallPercent(df, 0.5)
df_nona <- na.omit(df) #705 proteins
df_nona_qn <- normalizeQuantiles(df_nona)

pca_nona <- prcomp(t(df_nona_qn), scale = T)

summary(pca_nona)

fviz_pca_ind(pca_nona) #quick view 

principal_component_analysis(df_nona_qn)

pca_var <- pca_nona$sdev^2
pca_var_perc <- round(pca_var/sum(pca_var)*100, digits = 1) #percentage of each component

fviz_eig(pca_nona, addlabels = T) #screeplot

pca_results <- as.data.frame(pca_nona$x)
pca_pc1_pc2 <- pca_results %>% 
    dplyr::select(PC1, PC2) 

metadata_cleaned$expgrp <- as.factor(paste(metadata_cleaned$tissue, metadata_cleaned$intervention, metadata_cleaned$time, sep = "_"))

pca_pc1_pc2 <- pca_pc1_pc2 %>% 
    add_column(sample = rownames(pca_results),
               exp_group = metadata_cleaned$expgrp,
               tissue = metadata_cleaned$tissue, 
               intervention = metadata_cleaned$intervention,
               duration = metadata_cleaned$time,
               cryo_date = metadata_cleaned$cryostat,
               lcm_date = metadata_cleaned$lcm_date,
               sampleprep_occ = metadata_cleaned$sampleprep_occ,
               sampleprep_date = metadata_cleaned$sampleprep_date, 
               duration_frozen = metadata_cleaned$duration_frozen)

#####by tissue 
ggplot(pca_pc1_pc2, aes(x = PC1 , y = PC2, colour = tissue, label = sample)) +
    geom_point(size = 3) +
    xlab(paste("PC1 - ", pca_var_perc[1], "%", sep = "")) + 
    ylab(paste("PC2 - ", pca_var_perc[2], "%", sep = "")) + 
    ggtitle("caro_lcm_spect_PCA_noNA_qn") + 
    theme_minimal()
ggsave(file = "doc/caro_ex_lcm_spect_1sampleremoved_70f_matrix_qn_nona_PCA_byarea.jpeg")

#####by intervention and area 
ggplot(pca_pc1_pc2, aes(x = PC1 , y = PC2, colour = tissue, shape = intervention, label = sample)) +
    geom_point(size = 3) +
    scale_shape_manual(values = c("cc" = 1, "ecc" = 16, "sed" = 13)) +
    xlab(paste("PC1 - ", pca_var_perc[1], "%", sep = "")) + 
    ylab(paste("PC2 - ", pca_var_perc[2], "%", sep = "")) + 
    ggtitle("caro_lcm_spect_PCA_noNA_qn") + 
    theme_minimal()
ggsave(file = "doc/caro_ex_lcm_spect_1sampleremoved_70f_matrix_qn_nona_PCA_byint.jpeg")


#####by duration and area 
ggplot(pca_pc1_pc2, aes(x = PC1 , y = PC2, colour = tissue, shape = duration, label = sample)) +
    geom_point(size = 3) +
    scale_shape_manual(values = c("0" = 0, "5" = 15, "10" = 12)) +
    xlab(paste("PC1 - ", pca_var_perc[1], "%", sep = "")) + 
    ylab(paste("PC2 - ", pca_var_perc[2], "%", sep = "")) + 
    ggtitle("caro_lcm_spect_PCA_noNA_qn") + 
    theme_minimal()
ggsave(file = "doc/caro_ex_lcm_spect_1sampleremoved_70f_matrix_qn_nona_PCA_bytime.jpeg")

#####by expgrp
ggplot(pca_pc1_pc2, aes(x = PC1 , y = PC2, colour = exp_group, label = sample)) +
    geom_point(size = 3) +
    xlab(paste("PC1 - ", pca_var_perc[1], "%", sep = "")) + 
    ylab(paste("PC2 - ", pca_var_perc[2], "%", sep = "")) + 
    ggtitle("caro_lcm_spect_PCA_noNA_qn") + 
    theme_minimal()
ggsave(file = "doc/caro_ex_lcm_spect_1sampleremoved_70f_matrix_qn_nona_PCA_byexpgrp.jpeg")

#####by tissue and sampleprep_date
ggplot(pca_pc1_pc2, aes(x = PC1 , y = PC2, shape = sampleprep_date, colour = tissue, label = sample)) +
    geom_point(size = 3) +
    scale_shape_manual(values = c("1" = 2, "2" = 17)) +
    xlab(paste("PC1 - ", pca_var_perc[1], "%", sep = "")) + 
    ylab(paste("PC2 - ", pca_var_perc[2], "%", sep = "")) + 
    ggtitle("caro_lcm_spect_PCA_noNA_qn") + 
    theme_minimal()
ggsave(file = "doc/caro_ex_lcm_spect_1sampleremoved_70f_matrix_qn_nona_PCA_bysampleprepdate.jpeg")

#####by area and sampleprep_occ
ggplot(pca_pc1_pc2, aes(x = PC1 , y = PC2, shape = sampleprep_occ, colour = tissue, label = sample)) +
    geom_point(size = 3) +
    scale_shape_manual(values = c("1" = 1, "2" = 16, "3" = 10)) +
    xlab(paste("PC1 - ", pca_var_perc[1], "%", sep = "")) + 
    ylab(paste("PC2 - ", pca_var_perc[2], "%", sep = "")) + 
    ggtitle("caro_lcm_spect_PCA_noNA_qn") + 
    theme_minimal()
ggsave(file = "doc/caro_ex_lcm_spect_1sampleremoved_70f_matrix_qn_nona_PCA_bysampleprepocc.jpeg")

#####by cryo_date
ggplot(pca_pc1_pc2, aes(x = PC1 , y = PC2, colour = cryo_date, label = sample)) +
    geom_point(size = 3) +
    xlab(paste("PC1 - ", pca_var_perc[1], "%", sep = "")) + 
    ylab(paste("PC2 - ", pca_var_perc[2], "%", sep = "")) + 
    ggtitle("caro_lcm_spect_PCA_noNA_qn") + 
    theme_minimal()
ggsave(file = "doc/caro_ex_lcm_spect_1sampleremoved_70f_matrix_qn_nona_PCA_bycryo.jpeg")

#####by lcm_date
ggplot(pca_pc1_pc2, aes(x = PC1 , y = PC2, colour = lcm_date, label = sample)) +
    geom_point(size = 3) +
    xlab(paste("PC1 - ", pca_var_perc[1], "%", sep = "")) + 
    ylab(paste("PC2 - ", pca_var_perc[2], "%", sep = "")) + 
    ggtitle("caro_lcm_spect_PCA_noNA_qn") + 
    theme_minimal()
ggsave(file = "doc/caro_ex_lcm_spect_1sampleremoved_70f_matrix_qn_nona_PCA_bylcm.jpeg")


#####by duration_frozen
ggplot(pca_pc1_pc2, aes(x = PC1 , y = PC2, colour = duration_frozen, shape = sampleprep_date, label = sample)) +
    geom_point(size = 3) +
    scale_shape_manual(values = c("1" = 1, "2" = 19)) +
    xlab(paste("PC1 - ", pca_var_perc[1], "%", sep = "")) + 
    ylab(paste("PC2 - ", pca_var_perc[2], "%", sep = "")) + 
    ggtitle("caro_lcm_spect_PCA_noNA_qn") + 
    theme_minimal()
ggsave(file = "doc/caro_ex_lcm_spect_1sampleremoved_70f_matrix_qn_nona_PCA_bydurationfrozen.jpeg")
```


## Region specific visualization 
### MTJ 
```{r}
#df <- selectGrps(df_cleaned, metadata_cleaned$tissue, 0.5, n=1)
#if you first filter the data (whole matrix or in 1 group), you will end up having the same IDs. but if you divide the data to each region and then filter for 50%, we could then see unique IDs.

mtj <- metadata_cleaned %>% filter(metadata_cleaned$tissue == "MTJ")
df_mtj <- df_cleaned[ , trimws(mtj$sample_id)]

df_mtj_nona <- na.omit(df_mtj)

df_mtj50 <- selectOverallPercent(df_mtj, 0.5)
df_mtj50$median <- rowMedians(as.matrix(df_mtj50), na.rm = T)
df_mtj50 <- df_mtj50 %>% arrange(desc(median))
df_mtj50$rank = (1:2121) 
df_mtj50$protein <- row.names(df_mtj50)

ggplot(df_mtj50, aes(rank, median)) + 
  geom_line() +
    labs(title = "S-curve_allsamples_MTJ") + 
    ylab("Log10 protein intensities (median)") + 
    xlab("Abundance rank") + 
    theme_minimal()

muscle <- metadata_cleaned %>% filter(metadata_cleaned$tissue == "Muscle")
df_muscle <- df_cleaned[ , trimws(muscle$sample_id)]
df_muscle_nona <- na.omit(df_muscle)

df_muscle50 <- selectOverallPercent(df_muscle, 0.5)
df_muscle50$median <- rowMedians(as.matrix(df_muscle50), na.rm = T)
df_muscle50 <- df_muscle50 %>% arrange(desc(median))
df_muscle50$rank = (1:1826) 
df_muscle50$protein <- row.names(df_muscle50)

ggplot(df_muscle50, aes(rank, median)) + 
  geom_line() +
    labs(title = "S-curve_allsamples_muscle") + 
    ylab("Log10 protein intensities (median)") + 
    xlab("Abundance rank") + 
    theme_minimal()


tendon <- metadata_cleaned %>% filter(metadata_cleaned$tissue == "Tendon")
df_tendon <- df_cleaned[ , trimws(tendon$sample_id)]
df_tendon_nona <- na.omit(df_tendon)

df_tendon50 <- selectOverallPercent(df_tendon, 0.5)
df_tendon50$median <- rowMedians(as.matrix(df_tendon50), na.rm = T)
df_tendon50 <- df_tendon50 %>% arrange(desc(median))
df_tendon50$rank = (1:2015) 
df_tendon50$protein <- row.names(df_tendon50)

ggplot(df_tendon50, aes(rank, median)) + 
  geom_line() +
    labs(title = "S-curve_allsamples_tendon") + 
    ylab("Log10 protein intensities (median)") + 
    xlab("Abundance rank") + 
    theme_minimal()

###########################
#Venn diagram showing overlap between regions_filtered for no NA
overlap_nona <- list(MTJ = rownames(df_mtj_nona), 
                Tendon = rownames(df_tendon_nona),
                Muscle = rownames(df_muscle_nona))

ggVennDiagram(overlap_nona, label = "count", label_alpha = 0.8, set_size = 3) + 
    labs(title = "nona_ineachregion")


overlap_50f <- list(MTJ = df_mtj50$protein, 
                Tendon = df_tendon50$protein,
                Muscle = df_muscle50$protein)

ggVennDiagram(overlap_50f, label = "count", label_alpha = 0.8, set_size = 3) + 
    labs(title = "50f_ineachregion")


#common proteins
common_mtj_tendon <- intersect(df_mtj50$protein, df_tendon50$protein)
common_mtj_muscle <- intersect(df_mtj50$protein, df_muscle50$protein)
common_muscle_tendon <- intersect(df_tendon50$protein, df_muscle50$protein)

#exclusive in mtj 
exclusive_mtj <- setdiff(df_mtj50$protein, common_mtj_tendon)
exclusive_mtj <- setdiff(exclusive_mtj, common_mtj_muscle)

#exclusive in muscle 
exclusive_muscle <- setdiff(df_muscle50$protein, common_mtj_muscle)
exclusive_muscle <- setdiff(exclusive_muscle, common_muscle_tendon)

#exclusive in tendon 
exclusive_tendon <- setdiff(df_tendon50$protein, common_mtj_tendon)
exclusive_tendon <- setdiff(exclusive_tendon, common_muscle_tendon)
###########################
# MDS plot 
mds_mtj_70f_qn <- plotMDS(mtj_70f_qn)

toplot_mtj = data.frame(Dim1 = mds_mtj_70f_qn$x, Dim2 = mds_mtj_70f_qn$y, 
                    samples = mtj$sample_id, 
                    area = mtj$tissue, 
                    intervention = mtj$intervention, 
                    cryo_date = mtj$cryostat, 
                    lcm_date = mtj$lcm, 
                    sampleprep_date = mtj$sampleprep,
                    duration_frozen_slide = mtj$duration_frozen_slide, 
                    duration_frozen_lcmtissue = mtj$duration_frozen_microdis)

#####by intervention 
ggplot(toplot_mtj, aes(x = Dim1 , y = Dim2, shape = intervention, label = samples)) +
    geom_point(size = 3) +
    scale_shape_manual(values = c("cc" = 1, "ecc" = 13, "sed" = 19)) +
    xlab(paste("Leading logFC dim1 - 63%")) + 
    ylab(paste("Leading logFC dim2 - 7%")) + 
    ggtitle("df_1sampleremoved_70fNA_matrix_qn_MTJ_MDS") + 
    theme_minimal()
ggsave(file = "doc/caro_ex_lcm_spect_1sampleremoved_70f_matrix_qn_MTJ_MDS_byintervention.jpeg")

#####by sampleprep_date
ggplot(toplot_mtj, aes(x = Dim1 , y = Dim2, shape = sampleprep_date, label = samples)) +
    geom_point(size = 3) +
    scale_shape_manual(values = c("07/Nov" = 0, "09/Nov" = 15)) +
    xlab(paste("Leading logFC dim1 - 63%")) + 
    ylab(paste("Leading logFC dim2 - 7%")) + 
    ggtitle("df_1sampleremoved_70fNA_matrix_qn_MTJ_MDS") + 
    theme_minimal()
ggsave(file = "doc/caro_ex_lcm_spect_1sampleremoved_70f_matrix_qn_MTJ_MDS_bysampleprep.jpeg")

#####by cryo_date
ggplot(toplot_mtj, aes(x = Dim1 , y = Dim2, shape = cryo_date, colour = sampleprep_date, label = samples)) +
    geom_point(size = 3) +
    scale_shape_manual(values = c("02/Oct" = 1, "06/Oct" = 2, "11/Oct" = 3, "22/Sep" = 4, "22/Sep" = 5, "25/Oct" = 6, "25/Sep" = 7, "26/Oct" = 8, "27/Sep" = 9, "28/Oct" = 10, "28/Sep" = 11, "29/Sep" = 12)) +
    xlab(paste("Leading logFC dim1 - 63%")) + 
    ylab(paste("Leading logFC dim2 - 7%")) + 
    ggtitle("df_1sampleremoved_70fNA_matrix_qn_MDS") + 
    theme_minimal()
ggsave(file = "doc/caro_ex_lcm_spect_1sampleremoved_70f_matrix_qn_MDS_bycryo.jpeg")

#####by lcm_date
ggplot(toplot_mtj, aes(x = Dim1 , y = Dim2, shape = lcm_date, colour = sampleprep_date, label = samples)) +
    geom_point(size = 3) +
    scale_shape_manual(values = c("01/Nov" = 1, "02/Nov" = 2, "10/Oct" = 4, "16/Oct" = 6, "17/Oct" = 7, "18/Oct" = 8, "20/Oct" = 9, "23/Oct" = 10, "24/Oct" = 11, "26/Oct" = 12, "27/Oct" = 13, "30/Oct" = 14, "31/Oct" = 15)) +
    xlab(paste("Leading logFC dim1 - 63%")) + 
    ylab(paste("Leading logFC dim2 - 7%")) + 
    ggtitle("df_1sampleremoved_70fNA_matrix_qn_MTJ_MDS") + 
    theme_minimal()
ggsave(file = "doc/caro_ex_lcm_spect_1sampleremoved_70f_matrix_qn_MTJ_MDS_bylcm.jpeg")

#####by duration_frozen_days
ggplot(toplot_mtj, aes(x = Dim1 , y = Dim2, shape = duration_frozen_slide , colour = sampleprep_date , label = samples)) +
    geom_point(size = 3) +
    scale_shape_manual(values = c("1" = 1, "4" = 2, "5" = 3, "6" = 4, "7" = 5, "13" = 6, 
                       "18" = 7, "19" = 8, "20" = 9, "21" = 10, "25" = 11, "26" = 12, "27" = 13, 
                       "28" = 14, "31" = 15, "34" = 16, "35" = 17, "2" = 18)) +
    xlab(paste("Leading logFC dim1 - 63%")) + 
    ylab(paste("Leading logFC dim2 - 7%")) + 
    ggtitle("df_1sampleremoved_70fNA_matrix_qn_MTJ_MDS") + 
    theme_minimal()
ggsave(file = "doc/caro_ex_lcm_spect_1sampleremoved_70f_matrix_qn_MTJ_MDS_bydurationfrozenslide.jpeg")

#####by duration_frozen_days tissues microdissected
ggplot(toplot_mtj, aes(x = Dim1 , y = Dim2, shape = duration_frozen_lcmtissue, colour =  sampleprep_date, label = samples)) +
    geom_point(size = 3) +
    scale_shape_manual(values = c("1" = 1, "4" = 2, "5" = 3, "6" = 4, "7" = 5, "13" = 6, 
                       "18" = 7, "19" = 8, "20" = 9, "21" = 10, "25" = 11, "26" = 12, "27" = 13, 
                       "28" = 14, "31" = 15, "34" = 16, "35" = 17, "2" = 18)) +
    xlab(paste("Leading logFC dim1 - 63%")) + 
    ylab(paste("Leading logFC dim2 - 7%")) + 
    ggtitle("df_1sampleremoved_70fNA_matrix_qn_MTJ_MDS") + 
    theme_minimal()
#ggsave(file = "doc/caro_ex_lcm_spect_1sampleremoved_70f_matrix_qn_MTJ_MDS_bydurationfrozenlcmtissue.jpeg")

ggplot(toplot_mtj, aes(x = Dim1 , y = Dim2, shape = duration_frozen_lcmtissue, colour = sampleprep_date, label = samples)) +
    geom_point(size = 3) +
    scale_shape_manual(values = c("5" = 1, "6" = 2, "7" = 3, "8" = 4, "9" = 5, "10" = 6, "11" = 7, "12" = 8, "13" = 9, "14" = 10, "15" = 11, "17" = 12, "18" = 13, "21" = 14, "22" = 15, "23" = 16, "24" = 17, "28" = 23)) +
    xlab(paste("Leading logFC dim1 - 63%")) + 
    ylab(paste("Leading logFC dim2 - 7%")) + 
    ggtitle("df_1sampleremoved_70fNA_matrix_qn_MTJ_MDS") + 
    theme_minimal()
ggsave(file = "doc/caro_ex_lcm_spect_1sampleremoved_70f_matrix_qn_MTJ_MDS_bydurationfrozenlcmtissue.jpeg")

```




#Scurve 
this is a dynamic range plot of ranked protein abundance
```{r}

df_scurve <- df %>% #use the original df 
    mutate(genes = ifelse(genes == "", protein_names, genes)) %>%
    mutate_at(3:110, as.numeric) %>%
    mutate_all(~ifelse(is.nan(.), NA, .)) %>%
    mutate(genes = make.names(genes, unique = T), genes) %>% 
    column_to_rownames(var = "genes") %>% 
    dplyr::select(-c("protein_names")) %>%
    log10()

#str(df_scurve)

df_scurve <- df_70f_qn
df_scurve$median = rowMedians(as.matrix(df_scurve), na.rm = T)
df_scurve <- df_scurve %>% arrange(desc(median))
df_scurve$rank = (1:1784)

ggplot(df_scurve, aes(rank, median)) + 
  geom_line() +
    labs(title = "S-curve_allsamples") + 
    ylab("Log10 protein intensities (median)") + 
    xlab("Abundance rank") + 
    theme_minimal()

#the 10 most abundant proteins
row.names(df_scurve[1:10,])

#the 10 least abundant proteins
row.names(df_scurve[2309:2318,])

#ggsave(file = "doc/caro_ex_lcm_spect_allsamples_scurve.jpeg", width=10, height = 8)
```


###PCA only in MTJ 
here, i will check if there could be any clustering amoung the expgrpoup which combines tissue, intervention and int duration. 
```{r}
metadata_mtj <- metadata[metadata$tissue == "MTJ", ]
mtj_order = metadata_mtj$sample_id[order(metadata_mtj$intervention)]
mtj_order = trimws(mtj_order)

df_nona_qn_mtj = df_nona_qn[ ,mtj_order]


pca_nona_qn_mtj <- prcomp(t(df_nona_qn_mtj), scale = T)

summary(pca_nona_qn_mtj)
fviz_pca_ind(pca_nona_qn_mtj) #quick view 
plot(pca_nona_qn_mtj$x[,1], pca_nona_qn_mtj$x[,2]) #altenative 

pca_var <- pca_nona_qn_mtj$sdev^2
pca_var_perc <- round(pca_var/sum(pca_var)*100, digits = 1) #percentage of each component

fviz_eig(pca_nona_qn_mtj, addlabels = T) #screeplot

pca_results <- as.data.frame(pca_nona_qn_mtj$x)
pca_mtj_pc1_pc2 <- pca_results %>% 
    dplyr::select(PC1, PC2) 

pca_mtj_pc1_pc2 <- pca_mtj_pc1_pc2 %>% 
    add_column(sample = rownames(pca_results),
               exp_group = metadata_mtj$expgrp,
               tissue = metadata_mtj$tissue, 
               intervention = metadata_mtj$intervention,
               duration = metadata_mtj$time,
               cryo_date = metadata_mtj$cryostat_date,
               lcm_date = metadata_mtj$lcm_date,
               sampleprep_occ = metadata_mtj$sampleprep_occ,
               sampleprep_date = metadata_mtj$sampleprep_date)


#####by intervention
ggplot(pca_mtj_pc1_pc2, aes(x = PC1 , y = PC2, colour = intervention, label = sample)) +
    geom_point(size = 3) +
    #geom_text(fontface = "bold", show.legend = F, vjust = 0.5, hjust = -0.5, size = 3) + 
    xlab(paste("PC1 - ", pca_var_perc[1], "%", sep = "")) + 
    ylab(paste("PC2 - ", pca_var_perc[2], "%", sep = "")) + 
    #xlim(-50, 50) + ylim(-50, 50) +
    ggtitle("caro_lcm_spect_PCA_noNA_qn_MTJ") + 
    theme_minimal()
#ggsave(file = "../doc/caro_lcm_spect_noNA_qn_MTJ_PCA_byintervention.jpeg")


#####by duration and intervention 
ggplot(pca_mtj_pc1_pc2, aes(x = PC1 , y = PC2, colour = intervention, shape = duration, label = sample)) +
    geom_point(size = 3) +
    scale_shape_manual(values = c("0" = 0, "5" = 15, "10" = 12)) +
    #geom_text(fontface = "bold", show.legend = F, vjust = 0.5, hjust = -0.5, size = 3) + 
    xlab(paste("PC1 - ", pca_var_perc[1], "%", sep = "")) + 
    ylab(paste("PC2 - ", pca_var_perc[2], "%", sep = "")) + 
    #xlim(-50, 50) + ylim(-50, 50) +
    ggtitle("caro_lcm_spect_PCA_noNA_qn_MTJ") + 
    theme_minimal()
#ggsave(file = "../doc/caro_lcm_spect_noNA_qn_MTJ_PCA_byintervention_time.jpeg")

#####by expgrp
ggplot(pca_mtj_pc1_pc2, aes(x = PC1 , y = PC2, colour = exp_group, label = sample)) +
    geom_point(size = 3) +
    #geom_text(fontface = "bold", show.legend = F, vjust = 0.5, hjust = -0.5, size = 3) + 
    xlab(paste("PC1 - ", pca_var_perc[1], "%", sep = "")) + 
    ylab(paste("PC2 - ", pca_var_perc[2], "%", sep = "")) + 
    #xlim(-50, 50) + ylim(-50, 50) +
    ggtitle("caro_lcm_spect_PCA_noNA_qn_MTJ") + 
    theme_minimal()
#ggsave(file = "../doc/caro_lcm_spect_noNA_qn_MTJ_PCA_byexpgrp.jpeg")

#####by expgrp
ggplot(pca_mtj_pc1_pc2, aes(x = PC1 , y = PC2, shape = exp_group, label = sample)) +
    geom_point(size = 3) +
    scale_shape_manual(values = c("MTJ-cc-0" = 0, "MTJ-cc-10" = 15, "MTJ-cc-5" = 7,
                                  "MTJ-ecc-0" = 1, "MTJ-ecc-10" = 16, "MTJ-ecc-5" =10,
                                  "MTJ-sed-0" = 2, "MTJ-sed-10" = 17, "MTJ-sed-5" =6)) +
    #geom_text(fontface = "bold", show.legend = F, vjust = 0.5, hjust = -0.5, size = 3) + 
    xlab(paste("PC1 - ", pca_var_perc[1], "%", sep = "")) + 
    ylab(paste("PC2 - ", pca_var_perc[2], "%", sep = "")) + 
    #xlim(-50, 50) + ylim(-50, 50) +
    ggtitle("caro_lcm_spect_PCA_noNA_qn_MTJ") + 
    theme_minimal()
#ggsave(file = "../doc/caro_lcm_spect_noNA_qn_MTJ_PCA_byexpgrp.jpeg")


#####by tissue and sampleprep_date
ggplot(pca_mtj_pc1_pc2, aes(x = PC1 , y = PC2, shape = sampleprep_date, colour = intervention, label = sample)) +
    geom_point(size = 3) +
    scale_shape_manual(values = c("1" = 2, "2" = 17)) +
    xlab(paste("PC1 - ", pca_var_perc[1], "%", sep = "")) + 
    ylab(paste("PC2 - ", pca_var_perc[2], "%", sep = "")) + 
    ggtitle("caro_lcm_spect_PCA_noNA_qn_MTJ") + 
    theme_minimal()
#ggsave(file = "../doc/caro_lcm_spect_noNA_qn_MTJ_PCA_bysampleprepdate.jpeg")

```

