---
title: "2_data_vis_regions_hh"
output: html_document
date: "2024-02-21"
---
# load library, data, metadata
```{r}
source(here::here("R/library.R"))
load(here::here("data/df_cleaned.rda"))
load(here::here("data/metadata_cleaned.rda"))
```


# common/unique protein ID in each region
```{r}
#df <- selectGrps(df_cleaned, metadata_cleaned$tissue, 0.5, n=1)
#if you first filter the data (whole matrix or in 1 group), you will end up having the same IDs. but if you divide the data to each region and then filter for 50%, we could then see unique IDs.
mtj <- metadata_cleaned %>% filter(metadata_cleaned$tissue == "MTJ")
df_mtj <- df_cleaned[ , trimws(mtj$sample_id)]

df_mtj50 <- selectOverallPercent(df_mtj, 0.5)
df_mtj50$median <- rowMedians(as.matrix(df_mtj50), na.rm = T)
df_mtj50 <- df_mtj50 %>% arrange(desc(median))
df_mtj50$rank = (1:2121) 
df_mtj50$protein <- row.names(df_mtj50)

################
muscle <- metadata_cleaned %>% filter(metadata_cleaned$tissue == "Muscle")
df_muscle <- df_cleaned[ , trimws(muscle$sample_id)]

df_muscle50 <- selectOverallPercent(df_muscle, 0.5)
df_muscle50$median <- rowMedians(as.matrix(df_muscle50), na.rm = T)
df_muscle50 <- df_muscle50 %>% arrange(desc(median))
df_muscle50$rank = (1:1826) 
df_muscle50$protein <- row.names(df_muscle50)

###############
tendon <- metadata_cleaned %>% filter(metadata_cleaned$tissue == "Tendon")
df_tendon <- df_cleaned[ , trimws(tendon$sample_id)]

df_tendon50 <- selectOverallPercent(df_tendon, 0.5)
df_tendon50$median <- rowMedians(as.matrix(df_tendon50), na.rm = T)
df_tendon50 <- df_tendon50 %>% arrange(desc(median))
df_tendon50$rank = (1:2015) 
df_tendon50$protein <- row.names(df_tendon50)

###########################
#Venn diagram showing overlap between regions_filtered for 50% 
overlap_50f <- list(MTJ = df_mtj50$protein, 
                Tendon = df_tendon50$protein,
                Muscle = df_muscle50$protein)

ggVennDiagram(overlap_50f, label = "count", label_alpha = 0.8, set_size = 3) + 
    labs(title = "50f_ineachregion")


#common proteins
common_mtj_tendon <- intersect(df_mtj50$protein, df_tendon50$protein)
common_mtj_muscle <- intersect(df_mtj50$protein, df_muscle50$protein)
common_muscle_tendon <- intersect(df_tendon50$protein, df_muscle50$protein)

#exclusive in mtj 
exclusive_mtj <- setdiff(df_mtj50$protein, common_mtj_tendon)
exclusive_mtj <- setdiff(exclusive_mtj, common_mtj_muscle)

#exclusive in muscle 
exclusive_muscle <- setdiff(df_muscle50$protein, common_mtj_muscle)
exclusive_muscle <- setdiff(exclusive_muscle, common_muscle_tendon)

#exclusive in tendon 
exclusive_tendon <- setdiff(df_tendon50$protein, common_mtj_tendon)
exclusive_tendon <- setdiff(exclusive_tendon, common_muscle_tendon)

#Venn diagram showing overlap between regions_filtered for no NA
#df_mtj_nona <- na.omit(df_mtj)
#df_muscle_nona <- na.omit(df_muscle)
#df_tendon_nona <- na.omit(df_tendon)

#overlap_nona <- list(MTJ = rownames(df_mtj_nona), 
#                Tendon = rownames(df_tendon_nona),
#                Muscle = rownames(df_muscle_nona))

#ggVennDiagram(overlap_nona, label = "count", label_alpha = 0.8, set_size = 3) + 
#    labs(title = "nona_ineachregion")
```

# Scurve 
```{r}
# MTJ 
ggplot(df_mtj50, aes(rank, median)) + 
  geom_line() +
    geom_point(data = top_bottom, aes(color = "red"), size = 2) + 
  geom_text(data = top_bottom, aes(label = row.names(top_bottom)), vjust = 0.5, hjust = -0.5, size = 2.3) + 
    labs(title = "S-curve_allsamples_MTJ") + 
    ylab("Log2 protein intensities (median)") + 
    xlab("Abundance rank") + 
    theme_minimal()

mtj_top <- rownames(df_mtj50[1:10,])



# Muscle 
ggplot(df_muscle50, aes(rank, median)) + 
  geom_line() +
    labs(title = "S-curve_allsamples_muscle") + 
    ylab("Log2 protein intensities (median)") + 
    xlab("Abundance rank") + 
    
    theme_minimal()

# Tendon
ggplot(df_tendon50, aes(rank, median)) + 
  geom_line() +
    labs(title = "S-curve_allsamples_tendon") + 
    ylab("Log2 protein intensities (median)") + 
    xlab("Abundance rank") + 
    theme_minimal()


top_bottom <- c("H4c1", "Tuba1a", "Calm1.Calm2.Calm3", "Coq6", 
                "Camk2b", "Actc1", "Akt2","Pcdh18", "Cnot6", "Nrsn2")

top_bottom <- df_scurve[row.names(df_scurve) %in% top_bottom, ]


ggplot(df_scurve, aes(rank, median)) + 
  geom_line() +
  geom_point(data = top_bottom, aes(color = "red"), size = 2) + 
  geom_text(data = top_bottom, aes(label = row.names(top_bottom)), vjust = 0.5, hjust = -0.5, size = 2.3) + 
  labs(title = "S-curve_allsamples_6417_median") + 
  ylab("Log10 protein intensities (median)") + 
  xlab("Abundance rank") + 
  theme_minimal()
```




# proteins from literature
```{r}
# remove sample_ids that we are not so sure if they match the right tissue in PCA 
df_cleaned <- df_cleaned %>% dplyr::select(-"19", -"53", -"38")

metadata_cleaned <- metadata_cleaned %>%  filter(sample_id != "19", 
                                                 sample_id != "53", 
                                                 sample_id != "38")

# filter data for 50% in each group: 
df_cleaned_50f <- selectGrps(df_cleaned, metadata_cleaned$tissue, 0.5, n=3)

####can we merge so we can see exclusive proteins in each region... 

known_mtj_prot <- c("Xirp2", "Thbs4", "Ncam1", "Col22a1", "Fermt2", "Cd151", "Cilp", "Itga10", "Ogn", "Fmod", "Bgn", "Col1a2", "Col6a3", "Col6a2", "Col1a1", "Col6a1", "Col12a1", "Prelp", "Lam2", "Vdaxc1", "Vdac2", "Vdac3", "Mpz", "Vcl", "Plec", "Des", "Dsp", "Actn" , "Capz", "Flnc", "Fhl1", "Ttn", "Pdlim7", "Myot")

poi_indata <- df_cleaned[row.names(df_cleaned) %in% known_mtj_prot, ]

poi_indata_long <- poi_indata %>% 
    tibble::rownames_to_column("genes") %>% 
    tidyr::pivot_longer(
        cols = !genes,
        values_to = "intensities",
        names_to = "sample_id"
    ) %>% 
    dplyr::inner_join(metadata_cleaned) %>% 
    dplyr::select(genes, sample_id, intensities, tissue, intervention)

head(poi_indata_long)

ggplot(poi_indata_long, aes(x = tissue, y = intensities)) + 
  geom_violin(trim = FALSE) +  # Set trim to FALSE to display the entire violin plot
  geom_point(position = position_jitter(width = 0.2), size = 1.5, alpha = 0.5) +
  stat_summary(fun = median, geom = "crossbar", width = 0.5, color = "black") +
  labs(x = "Tissue", y = "log2 protein intensities", title = "known_mtj_proteins") +
  theme_minimal() + 
  facet_wrap(~ genes, scales = "free")

################
known_tendon_prot <- c("Serpina1", "Dcn", "Fbn1", 
"Fn1", "Col1a1", "Col1a2", "Col3a1", "Col4a1", "Col5a1", "Col5a2", 
"Col6a1", "Col6a2", "Col6a3", "Col11a1", "Col11a2", "Col2a1", "Col12a1", "Comp")

poi_indata <- df_cleaned[row.names(df_cleaned) %in% known_tendon_prot, ]

poi_indata_long <- poi_indata %>% 
    tibble::rownames_to_column("genes") %>% 
    tidyr::pivot_longer(
        cols = !genes,
        values_to = "intensities",
        names_to = "sample_id"
    ) %>% 
    dplyr::inner_join(metadata_cleaned) %>% 
    dplyr::select(genes, sample_id, intensities, tissue, intervention)

head(poi_indata_long)

ggplot(poi_indata_long, aes(x = tissue, y = intensities)) + 
  geom_violin(trim = FALSE) +  # Set trim to FALSE to display the entire violin plot
  geom_point(position = position_jitter(width = 0.2), size = 2, alpha = 0.5) +
  stat_summary(fun = median, geom = "crossbar", width = 0.5, color = "black") +
  labs(x = "Tissue", y = "log2 protein intensities", title = "known_tendon_proteins") +
  theme_minimal() + 
  facet_wrap(~ genes, scales = "free")

```



#NA filter whole matrix and normalize
raw data matrix is saved as Rda file. Similarly, data is saved separately after NA filtering.
```{r}
df_70f <- selectOverallPercent(df_cleaned, 0.7) #1784 proteins << original df was 2318

df_70f_qn <- normalizeQuantiles(df_70f)

usethis::use_data(df_70f_qn, overwrite = T)
write.csv(df_70f_qn, file = "data/caro_ex_lcm_spect_1sampleremoved_sorted_70f_qn.csv", row.names = T, quote = F)
```


#data visualization
```{r}
metadata.grp.col <- c("#F8766D", "#00BA38", "#619CFF")
grp.col <- metadata.grp.col[match(metadata_cleaned$tissue, unique(metadata_cleaned$tissue))]

boxplot(df_70f_qn, main = "caro_spect_1sampleremoved_NA70f_qn", ylab = "log2 intensities", col = grp.col)
legend("topleft", inset = .02,
       c("MTJ","Muscle","Tendon"), 
       fill = metadata.grp.col, 
       horiz = TRUE, cex = 0.8)
```

# MDS plot 
***MDS shows a region-specific clustering*** There is no clustering based on intervention, duration of intervention, MS sampleprep date, cryostat date, lcm date or duration of freezing. 
```{r}
mds_70f_qn <- plotMDS(df_70f_qn)

toplot = data.frame(Dim1 = mds_70f_qn$x, Dim2 = mds_70f_qn$y, 
                    samples = metadata_cleaned$sample_id, 
                    area = metadata_cleaned$tissue, 
                    intervention = metadata_cleaned$intervention, 
                    cryo_date = metadata_cleaned$cryostat, 
                    lcm_date = metadata_cleaned$lcm, 
                    sampleprep_date = metadata_cleaned$sampleprep_date,
                    duration_frozen_slide = metadata_cleaned$duration_frozen_slide, 
                    duration_frozen_lcmtissue = metadata_cleaned$duration_frozen_microdis)

#####by area 
ggplot(toplot, aes(x = Dim1 , y = Dim2, colour = area, label = samples)) +
    geom_point(size = 3) +
    xlab(paste("Leading logFC dim1 - 63%")) + 
    ylab(paste("Leading logFC dim2 - 7%")) + 
    ggtitle("df_1sampleremoved_70fNA_matrix_qn_MDS") + 
    theme_minimal()
ggsave(file = "doc/caro_ex_lcm_spect_1sampleremoved_70f_matrix_qn_MDS_byarea.jpeg")

#####by intervention 
ggplot(toplot, aes(x = Dim1 , y = Dim2, colour = area, shape = intervention, label = samples)) +
    geom_point(size = 3) +
    scale_shape_manual(values = c("cc" = 1, "ecc" = 13, "sed" = 19)) +
    xlab(paste("Leading logFC dim1 - 63%")) + 
    ylab(paste("Leading logFC dim2 - 7%")) + 
    ggtitle("df_1sampleremoved_70fNA_matrix_qn_MDS") + 
    theme_minimal()
ggsave(file = "doc/caro_ex_lcm_spect_1sampleremoved_70f_matrix_qn_MDS_byintervention.jpeg")

#####by sampleprep_date
ggplot(toplot, aes(x = Dim1 , y = Dim2, colour = area, shape = sampleprep_date, label = samples)) +
    geom_point(size = 3) +
    scale_shape_manual(values = c("1" = 0, "2" = 15)) +
    xlab(paste("Leading logFC dim1 - 63%")) + 
    ylab(paste("Leading logFC dim2 - 7%")) + 
    ggtitle("df_1sampleremoved_70fNA_matrix_qn_MDS") + 
    theme_minimal()
ggsave(file = "doc/caro_ex_lcm_spect_1sampleremoved_70f_matrix_qn_MDS_bysampleprep.jpeg")

#####by cryo_date
ggplot(toplot, aes(x = Dim1 , y = Dim2, shape = cryo_date, colour = sampleprep_date, label = samples)) +
    geom_point(size = 3) +
    scale_shape_manual(values = c("02/Oct" = 1, "06/Oct" = 2, "11/Oct" = 3, "22/Sep" = 4, "22/Sep" = 5, "25/Oct" = 6, "25/Sep" = 7, "26/Oct" = 8, "27/Sep" = 9, "28/Oct" = 10, "28/Sep" = 11, "29/Sep" = 12)) +
    xlab(paste("Leading logFC dim1 - 63%")) + 
    ylab(paste("Leading logFC dim2 - 7%")) + 
    ggtitle("df_1sampleremoved_70fNA_matrix_qn_MDS") + 
    theme_minimal()
ggsave(file = "doc/caro_ex_lcm_spect_1sampleremoved_70f_matrix_qn_MDS_bycryo.jpeg")

#####by lcm_date
ggplot(toplot, aes(x = Dim1 , y = Dim2, shape = lcm_date, colour = sampleprep_date, label = samples)) +
    geom_point(size = 3) +
    scale_shape_manual(values = c("01/Nov" = 1, "02/Nov" = 2, "09/Oct" = 3, "10/Oct" = 4, "12/Oct" = 5, "16/Oct" = 6, "17/Oct" = 7, "18/Oct" = 8, "20/Oct" = 9, "23/Oct" = 10, "24/Oct" = 11, "26/Oct" = 12, "27/Oct" = 13, "30/Oct" = 14, "31/Oct" = 15)) +
    xlab(paste("Leading logFC dim1 - 63%")) + 
    ylab(paste("Leading logFC dim2 - 7%")) + 
    ggtitle("df_1sampleremoved_70fNA_matrix_qn_MDS") + 
    theme_minimal()
ggsave(file = "doc/caro_ex_lcm_spect_1sampleremoved_70f_matrix_qn_MDS_bylcm.jpeg")


#####by duration_frozen_days slides
ggplot(toplot, aes(x = Dim1 , y = Dim2, shape = sampleprep_date, colour = duration_frozen_slide, label = samples)) +
    geom_point(size = 3) +
    scale_shape_manual(values = c("1" = 1, "2" = 19)) +
    xlab(paste("Leading logFC dim1 - 63%")) + 
    ylab(paste("Leading logFC dim2 - 7%")) + 
    ggtitle("df_1sampleremoved_70fNA_matrix_qn_MDS") + 
    theme_minimal()
ggsave(file = "doc/caro_ex_lcm_spect_1sampleremoved_70f_matrix_qn_MDS_bydurationfrozenslide.jpeg")


#####by duration_frozen_days tissues microdissected
ggplot(toplot, aes(x = Dim1 , y = Dim2, shape = sampleprep_date, colour = duration_frozen_lcmtissue, label = samples)) +
    geom_point(size = 3) +
    scale_shape_manual(values = c("1" = 1, "2" = 19)) +
    xlab(paste("Leading logFC dim1 - 63%")) + 
    ylab(paste("Leading logFC dim2 - 7%")) + 
    ggtitle("df_1sampleremoved_70fNA_matrix_qn_MDS") + 
    theme_minimal()
ggsave(file = "doc/caro_ex_lcm_spect_1sampleremoved_70f_matrix_qn_MDS_bydurationfrozenlcmtissue.jpeg")
```

# PCA using 100%filtered data (na.omit)
Here; all NAs are omitted, resulting **810 proteins** in total. Afterwards, data is quantile normalized.

***PCA shows a region-specific clustering*** There is no clustering based on intervention, duration of intervention, MS sampleprep date, cryostat date, lcm date or duration of freezing. 
```{r}

principal_component_analysis <-
  function(.data, our_variable) {
    data_pca <- t(.data) |>
      as.matrix()
    pca_object <- stats::prcomp(data_pca, scale = T)
    data_pca <- pca_object$x
    data_pca <- as.data.frame(data_pca)
    data_pca <- data_pca |>
      dplyr::select(PC1, PC2) |>
      tibble::add_column(our_variable = our_variable)
    pca_plot <- ggplot(data_pca,
                       aes(
                         x = PC1,
                         y = PC2,
                         color = our_variable,
                         names = sample_id
                       )) +
      geom_point() +
      labs(colour = deparse(substitute(our_variable)))+
      theme_minimal() +
      theme(plot.title = element_text(size = 16, face = "bold")) +
      theme(plot.title = element_text(hjust = 0.5)) +
      theme(legend.title = element_text(size = 12, face = "bold")) +
      scale_color_viridis_d()
    return(plotly::ggplotly(pca_plot))
  }



df_50f <- selectOverallPercent(df, 0.5)
df_nona <- na.omit(df) #705 proteins
df_nona_qn <- normalizeQuantiles(df_nona)

pca_nona <- prcomp(t(df_nona_qn), scale = T)

summary(pca_nona)

fviz_pca_ind(pca_nona) #quick view 

principal_component_analysis(df_nona_qn)

pca_var <- pca_nona$sdev^2
pca_var_perc <- round(pca_var/sum(pca_var)*100, digits = 1) #percentage of each component

fviz_eig(pca_nona, addlabels = T) #screeplot

pca_results <- as.data.frame(pca_nona$x)
pca_pc1_pc2 <- pca_results %>% 
    dplyr::select(PC1, PC2) 

metadata_cleaned$expgrp <- as.factor(paste(metadata_cleaned$tissue, metadata_cleaned$intervention, metadata_cleaned$time, sep = "_"))

pca_pc1_pc2 <- pca_pc1_pc2 %>% 
    add_column(sample = rownames(pca_results),
               exp_group = metadata_cleaned$expgrp,
               tissue = metadata_cleaned$tissue, 
               intervention = metadata_cleaned$intervention,
               duration = metadata_cleaned$time,
               cryo_date = metadata_cleaned$cryostat,
               lcm_date = metadata_cleaned$lcm_date,
               sampleprep_occ = metadata_cleaned$sampleprep_occ,
               sampleprep_date = metadata_cleaned$sampleprep_date, 
               duration_frozen = metadata_cleaned$duration_frozen)

#####by tissue 
ggplot(pca_pc1_pc2, aes(x = PC1 , y = PC2, colour = tissue, label = sample)) +
    geom_point(size = 3) +
    xlab(paste("PC1 - ", pca_var_perc[1], "%", sep = "")) + 
    ylab(paste("PC2 - ", pca_var_perc[2], "%", sep = "")) + 
    ggtitle("caro_lcm_spect_PCA_noNA_qn") + 
    theme_minimal()
ggsave(file = "doc/caro_ex_lcm_spect_1sampleremoved_70f_matrix_qn_nona_PCA_byarea.jpeg")

#####by intervention and area 
ggplot(pca_pc1_pc2, aes(x = PC1 , y = PC2, colour = tissue, shape = intervention, label = sample)) +
    geom_point(size = 3) +
    scale_shape_manual(values = c("cc" = 1, "ecc" = 16, "sed" = 13)) +
    xlab(paste("PC1 - ", pca_var_perc[1], "%", sep = "")) + 
    ylab(paste("PC2 - ", pca_var_perc[2], "%", sep = "")) + 
    ggtitle("caro_lcm_spect_PCA_noNA_qn") + 
    theme_minimal()
ggsave(file = "doc/caro_ex_lcm_spect_1sampleremoved_70f_matrix_qn_nona_PCA_byint.jpeg")


#####by duration and area 
ggplot(pca_pc1_pc2, aes(x = PC1 , y = PC2, colour = tissue, shape = duration, label = sample)) +
    geom_point(size = 3) +
    scale_shape_manual(values = c("0" = 0, "5" = 15, "10" = 12)) +
    xlab(paste("PC1 - ", pca_var_perc[1], "%", sep = "")) + 
    ylab(paste("PC2 - ", pca_var_perc[2], "%", sep = "")) + 
    ggtitle("caro_lcm_spect_PCA_noNA_qn") + 
    theme_minimal()
ggsave(file = "doc/caro_ex_lcm_spect_1sampleremoved_70f_matrix_qn_nona_PCA_bytime.jpeg")

#####by expgrp
ggplot(pca_pc1_pc2, aes(x = PC1 , y = PC2, colour = exp_group, label = sample)) +
    geom_point(size = 3) +
    xlab(paste("PC1 - ", pca_var_perc[1], "%", sep = "")) + 
    ylab(paste("PC2 - ", pca_var_perc[2], "%", sep = "")) + 
    ggtitle("caro_lcm_spect_PCA_noNA_qn") + 
    theme_minimal()
ggsave(file = "doc/caro_ex_lcm_spect_1sampleremoved_70f_matrix_qn_nona_PCA_byexpgrp.jpeg")

#####by tissue and sampleprep_date
ggplot(pca_pc1_pc2, aes(x = PC1 , y = PC2, shape = sampleprep_date, colour = tissue, label = sample)) +
    geom_point(size = 3) +
    scale_shape_manual(values = c("1" = 2, "2" = 17)) +
    xlab(paste("PC1 - ", pca_var_perc[1], "%", sep = "")) + 
    ylab(paste("PC2 - ", pca_var_perc[2], "%", sep = "")) + 
    ggtitle("caro_lcm_spect_PCA_noNA_qn") + 
    theme_minimal()
ggsave(file = "doc/caro_ex_lcm_spect_1sampleremoved_70f_matrix_qn_nona_PCA_bysampleprepdate.jpeg")

#####by area and sampleprep_occ
ggplot(pca_pc1_pc2, aes(x = PC1 , y = PC2, shape = sampleprep_occ, colour = tissue, label = sample)) +
    geom_point(size = 3) +
    scale_shape_manual(values = c("1" = 1, "2" = 16, "3" = 10)) +
    xlab(paste("PC1 - ", pca_var_perc[1], "%", sep = "")) + 
    ylab(paste("PC2 - ", pca_var_perc[2], "%", sep = "")) + 
    ggtitle("caro_lcm_spect_PCA_noNA_qn") + 
    theme_minimal()
ggsave(file = "doc/caro_ex_lcm_spect_1sampleremoved_70f_matrix_qn_nona_PCA_bysampleprepocc.jpeg")

#####by cryo_date
ggplot(pca_pc1_pc2, aes(x = PC1 , y = PC2, colour = cryo_date, label = sample)) +
    geom_point(size = 3) +
    xlab(paste("PC1 - ", pca_var_perc[1], "%", sep = "")) + 
    ylab(paste("PC2 - ", pca_var_perc[2], "%", sep = "")) + 
    ggtitle("caro_lcm_spect_PCA_noNA_qn") + 
    theme_minimal()
ggsave(file = "doc/caro_ex_lcm_spect_1sampleremoved_70f_matrix_qn_nona_PCA_bycryo.jpeg")

#####by lcm_date
ggplot(pca_pc1_pc2, aes(x = PC1 , y = PC2, colour = lcm_date, label = sample)) +
    geom_point(size = 3) +
    xlab(paste("PC1 - ", pca_var_perc[1], "%", sep = "")) + 
    ylab(paste("PC2 - ", pca_var_perc[2], "%", sep = "")) + 
    ggtitle("caro_lcm_spect_PCA_noNA_qn") + 
    theme_minimal()
ggsave(file = "doc/caro_ex_lcm_spect_1sampleremoved_70f_matrix_qn_nona_PCA_bylcm.jpeg")


#####by duration_frozen
ggplot(pca_pc1_pc2, aes(x = PC1 , y = PC2, colour = duration_frozen, shape = sampleprep_date, label = sample)) +
    geom_point(size = 3) +
    scale_shape_manual(values = c("1" = 1, "2" = 19)) +
    xlab(paste("PC1 - ", pca_var_perc[1], "%", sep = "")) + 
    ylab(paste("PC2 - ", pca_var_perc[2], "%", sep = "")) + 
    ggtitle("caro_lcm_spect_PCA_noNA_qn") + 
    theme_minimal()
ggsave(file = "doc/caro_ex_lcm_spect_1sampleremoved_70f_matrix_qn_nona_PCA_bydurationfrozen.jpeg")
```


## Region specific visualization 
### MTJ 
```{r}

###########################
# MDS plot 
mds_mtj_70f_qn <- plotMDS(mtj_70f_qn)

toplot_mtj = data.frame(Dim1 = mds_mtj_70f_qn$x, Dim2 = mds_mtj_70f_qn$y, 
                    samples = mtj$sample_id, 
                    area = mtj$tissue, 
                    intervention = mtj$intervention, 
                    cryo_date = mtj$cryostat, 
                    lcm_date = mtj$lcm, 
                    sampleprep_date = mtj$sampleprep,
                    duration_frozen_slide = mtj$duration_frozen_slide, 
                    duration_frozen_lcmtissue = mtj$duration_frozen_microdis)

#####by intervention 
ggplot(toplot_mtj, aes(x = Dim1 , y = Dim2, shape = intervention, label = samples)) +
    geom_point(size = 3) +
    scale_shape_manual(values = c("cc" = 1, "ecc" = 13, "sed" = 19)) +
    xlab(paste("Leading logFC dim1 - 63%")) + 
    ylab(paste("Leading logFC dim2 - 7%")) + 
    ggtitle("df_1sampleremoved_70fNA_matrix_qn_MTJ_MDS") + 
    theme_minimal()
ggsave(file = "doc/caro_ex_lcm_spect_1sampleremoved_70f_matrix_qn_MTJ_MDS_byintervention.jpeg")

#####by sampleprep_date
ggplot(toplot_mtj, aes(x = Dim1 , y = Dim2, shape = sampleprep_date, label = samples)) +
    geom_point(size = 3) +
    scale_shape_manual(values = c("07/Nov" = 0, "09/Nov" = 15)) +
    xlab(paste("Leading logFC dim1 - 63%")) + 
    ylab(paste("Leading logFC dim2 - 7%")) + 
    ggtitle("df_1sampleremoved_70fNA_matrix_qn_MTJ_MDS") + 
    theme_minimal()
ggsave(file = "doc/caro_ex_lcm_spect_1sampleremoved_70f_matrix_qn_MTJ_MDS_bysampleprep.jpeg")

#####by cryo_date
ggplot(toplot_mtj, aes(x = Dim1 , y = Dim2, shape = cryo_date, colour = sampleprep_date, label = samples)) +
    geom_point(size = 3) +
    scale_shape_manual(values = c("02/Oct" = 1, "06/Oct" = 2, "11/Oct" = 3, "22/Sep" = 4, "22/Sep" = 5, "25/Oct" = 6, "25/Sep" = 7, "26/Oct" = 8, "27/Sep" = 9, "28/Oct" = 10, "28/Sep" = 11, "29/Sep" = 12)) +
    xlab(paste("Leading logFC dim1 - 63%")) + 
    ylab(paste("Leading logFC dim2 - 7%")) + 
    ggtitle("df_1sampleremoved_70fNA_matrix_qn_MDS") + 
    theme_minimal()
ggsave(file = "doc/caro_ex_lcm_spect_1sampleremoved_70f_matrix_qn_MDS_bycryo.jpeg")

#####by lcm_date
ggplot(toplot_mtj, aes(x = Dim1 , y = Dim2, shape = lcm_date, colour = sampleprep_date, label = samples)) +
    geom_point(size = 3) +
    scale_shape_manual(values = c("01/Nov" = 1, "02/Nov" = 2, "10/Oct" = 4, "16/Oct" = 6, "17/Oct" = 7, "18/Oct" = 8, "20/Oct" = 9, "23/Oct" = 10, "24/Oct" = 11, "26/Oct" = 12, "27/Oct" = 13, "30/Oct" = 14, "31/Oct" = 15)) +
    xlab(paste("Leading logFC dim1 - 63%")) + 
    ylab(paste("Leading logFC dim2 - 7%")) + 
    ggtitle("df_1sampleremoved_70fNA_matrix_qn_MTJ_MDS") + 
    theme_minimal()
ggsave(file = "doc/caro_ex_lcm_spect_1sampleremoved_70f_matrix_qn_MTJ_MDS_bylcm.jpeg")

#####by duration_frozen_days
ggplot(toplot_mtj, aes(x = Dim1 , y = Dim2, shape = duration_frozen_slide , colour = sampleprep_date , label = samples)) +
    geom_point(size = 3) +
    scale_shape_manual(values = c("1" = 1, "4" = 2, "5" = 3, "6" = 4, "7" = 5, "13" = 6, 
                       "18" = 7, "19" = 8, "20" = 9, "21" = 10, "25" = 11, "26" = 12, "27" = 13, 
                       "28" = 14, "31" = 15, "34" = 16, "35" = 17, "2" = 18)) +
    xlab(paste("Leading logFC dim1 - 63%")) + 
    ylab(paste("Leading logFC dim2 - 7%")) + 
    ggtitle("df_1sampleremoved_70fNA_matrix_qn_MTJ_MDS") + 
    theme_minimal()
ggsave(file = "doc/caro_ex_lcm_spect_1sampleremoved_70f_matrix_qn_MTJ_MDS_bydurationfrozenslide.jpeg")

#####by duration_frozen_days tissues microdissected
ggplot(toplot_mtj, aes(x = Dim1 , y = Dim2, shape = duration_frozen_lcmtissue, colour =  sampleprep_date, label = samples)) +
    geom_point(size = 3) +
    scale_shape_manual(values = c("1" = 1, "4" = 2, "5" = 3, "6" = 4, "7" = 5, "13" = 6, 
                       "18" = 7, "19" = 8, "20" = 9, "21" = 10, "25" = 11, "26" = 12, "27" = 13, 
                       "28" = 14, "31" = 15, "34" = 16, "35" = 17, "2" = 18)) +
    xlab(paste("Leading logFC dim1 - 63%")) + 
    ylab(paste("Leading logFC dim2 - 7%")) + 
    ggtitle("df_1sampleremoved_70fNA_matrix_qn_MTJ_MDS") + 
    theme_minimal()
#ggsave(file = "doc/caro_ex_lcm_spect_1sampleremoved_70f_matrix_qn_MTJ_MDS_bydurationfrozenlcmtissue.jpeg")

ggplot(toplot_mtj, aes(x = Dim1 , y = Dim2, shape = duration_frozen_lcmtissue, colour = sampleprep_date, label = samples)) +
    geom_point(size = 3) +
    scale_shape_manual(values = c("5" = 1, "6" = 2, "7" = 3, "8" = 4, "9" = 5, "10" = 6, "11" = 7, "12" = 8, "13" = 9, "14" = 10, "15" = 11, "17" = 12, "18" = 13, "21" = 14, "22" = 15, "23" = 16, "24" = 17, "28" = 23)) +
    xlab(paste("Leading logFC dim1 - 63%")) + 
    ylab(paste("Leading logFC dim2 - 7%")) + 
    ggtitle("df_1sampleremoved_70fNA_matrix_qn_MTJ_MDS") + 
    theme_minimal()
ggsave(file = "doc/caro_ex_lcm_spect_1sampleremoved_70f_matrix_qn_MTJ_MDS_bydurationfrozenlcmtissue.jpeg")

```






###PCA only in MTJ 
here, i will check if there could be any clustering amoung the expgrpoup which combines tissue, intervention and int duration. 
```{r}
metadata_mtj <- metadata[metadata$tissue == "MTJ", ]
mtj_order = metadata_mtj$sample_id[order(metadata_mtj$intervention)]
mtj_order = trimws(mtj_order)

df_nona_qn_mtj = df_nona_qn[ ,mtj_order]


pca_nona_qn_mtj <- prcomp(t(df_nona_qn_mtj), scale = T)

summary(pca_nona_qn_mtj)
fviz_pca_ind(pca_nona_qn_mtj) #quick view 
plot(pca_nona_qn_mtj$x[,1], pca_nona_qn_mtj$x[,2]) #altenative 

pca_var <- pca_nona_qn_mtj$sdev^2
pca_var_perc <- round(pca_var/sum(pca_var)*100, digits = 1) #percentage of each component

fviz_eig(pca_nona_qn_mtj, addlabels = T) #screeplot

pca_results <- as.data.frame(pca_nona_qn_mtj$x)
pca_mtj_pc1_pc2 <- pca_results %>% 
    dplyr::select(PC1, PC2) 

pca_mtj_pc1_pc2 <- pca_mtj_pc1_pc2 %>% 
    add_column(sample = rownames(pca_results),
               exp_group = metadata_mtj$expgrp,
               tissue = metadata_mtj$tissue, 
               intervention = metadata_mtj$intervention,
               duration = metadata_mtj$time,
               cryo_date = metadata_mtj$cryostat_date,
               lcm_date = metadata_mtj$lcm_date,
               sampleprep_occ = metadata_mtj$sampleprep_occ,
               sampleprep_date = metadata_mtj$sampleprep_date)


#####by intervention
ggplot(pca_mtj_pc1_pc2, aes(x = PC1 , y = PC2, colour = intervention, label = sample)) +
    geom_point(size = 3) +
    #geom_text(fontface = "bold", show.legend = F, vjust = 0.5, hjust = -0.5, size = 3) + 
    xlab(paste("PC1 - ", pca_var_perc[1], "%", sep = "")) + 
    ylab(paste("PC2 - ", pca_var_perc[2], "%", sep = "")) + 
    #xlim(-50, 50) + ylim(-50, 50) +
    ggtitle("caro_lcm_spect_PCA_noNA_qn_MTJ") + 
    theme_minimal()
#ggsave(file = "../doc/caro_lcm_spect_noNA_qn_MTJ_PCA_byintervention.jpeg")


#####by duration and intervention 
ggplot(pca_mtj_pc1_pc2, aes(x = PC1 , y = PC2, colour = intervention, shape = duration, label = sample)) +
    geom_point(size = 3) +
    scale_shape_manual(values = c("0" = 0, "5" = 15, "10" = 12)) +
    #geom_text(fontface = "bold", show.legend = F, vjust = 0.5, hjust = -0.5, size = 3) + 
    xlab(paste("PC1 - ", pca_var_perc[1], "%", sep = "")) + 
    ylab(paste("PC2 - ", pca_var_perc[2], "%", sep = "")) + 
    #xlim(-50, 50) + ylim(-50, 50) +
    ggtitle("caro_lcm_spect_PCA_noNA_qn_MTJ") + 
    theme_minimal()
#ggsave(file = "../doc/caro_lcm_spect_noNA_qn_MTJ_PCA_byintervention_time.jpeg")

#####by expgrp
ggplot(pca_mtj_pc1_pc2, aes(x = PC1 , y = PC2, colour = exp_group, label = sample)) +
    geom_point(size = 3) +
    #geom_text(fontface = "bold", show.legend = F, vjust = 0.5, hjust = -0.5, size = 3) + 
    xlab(paste("PC1 - ", pca_var_perc[1], "%", sep = "")) + 
    ylab(paste("PC2 - ", pca_var_perc[2], "%", sep = "")) + 
    #xlim(-50, 50) + ylim(-50, 50) +
    ggtitle("caro_lcm_spect_PCA_noNA_qn_MTJ") + 
    theme_minimal()
#ggsave(file = "../doc/caro_lcm_spect_noNA_qn_MTJ_PCA_byexpgrp.jpeg")

#####by expgrp
ggplot(pca_mtj_pc1_pc2, aes(x = PC1 , y = PC2, shape = exp_group, label = sample)) +
    geom_point(size = 3) +
    scale_shape_manual(values = c("MTJ-cc-0" = 0, "MTJ-cc-10" = 15, "MTJ-cc-5" = 7,
                                  "MTJ-ecc-0" = 1, "MTJ-ecc-10" = 16, "MTJ-ecc-5" =10,
                                  "MTJ-sed-0" = 2, "MTJ-sed-10" = 17, "MTJ-sed-5" =6)) +
    #geom_text(fontface = "bold", show.legend = F, vjust = 0.5, hjust = -0.5, size = 3) + 
    xlab(paste("PC1 - ", pca_var_perc[1], "%", sep = "")) + 
    ylab(paste("PC2 - ", pca_var_perc[2], "%", sep = "")) + 
    #xlim(-50, 50) + ylim(-50, 50) +
    ggtitle("caro_lcm_spect_PCA_noNA_qn_MTJ") + 
    theme_minimal()
#ggsave(file = "../doc/caro_lcm_spect_noNA_qn_MTJ_PCA_byexpgrp.jpeg")


#####by tissue and sampleprep_date
ggplot(pca_mtj_pc1_pc2, aes(x = PC1 , y = PC2, shape = sampleprep_date, colour = intervention, label = sample)) +
    geom_point(size = 3) +
    scale_shape_manual(values = c("1" = 2, "2" = 17)) +
    xlab(paste("PC1 - ", pca_var_perc[1], "%", sep = "")) + 
    ylab(paste("PC2 - ", pca_var_perc[2], "%", sep = "")) + 
    ggtitle("caro_lcm_spect_PCA_noNA_qn_MTJ") + 
    theme_minimal()
#ggsave(file = "../doc/caro_lcm_spect_noNA_qn_MTJ_PCA_bysampleprepdate.jpeg")

```
